---
title: "CATS case study illustration: MI approaches for incomplete three-level data with time-varying cluster-memberships"
author: "Rushani Wijesuriya"
date:  June 24, 2021
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CATS case study illustration: MI approaches for incomplete three-level data with time-varying cluster-memberships}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides an illustration of all MI approaches  evaluated in the paper:

>Wijesuriya R., Moreno-Betancur M., Carlin J., De Silva A., and Lee K.: "Multiple imputation approaches for handling incomplete three-level data with time-varying cluster-memberships". 

## Loading and looking at the example dataset

This illustration uses  data stored [here](https://github.com/rushwije/CATS_illustration_cross-classfied) that were simulated mimicking the dependencies in the Childhood to Adolescence Transition Study (CATS). Missing values were generated in the exposure as explained in detail in the paper, and also in the outcome and the auxiliary variable similar to the CATS with missing values assigned completely at random for simplicity. The R script for generating the example dataset can be found [here](https://github.com/rushwije/CATS_illustration_cross-classfied/blob/main/Data%20simulation.R). 

The dataset contains the following variables:

* `uniqueid`: Child's ID
* `c1dage`: Child’s age (wave 1) (years) 
* `cdgender`: Child’s sex 
* `SES`: Standardized SES measured by the SEIFA IRSAD (wave 1)
* `tscore_W1`: Teacher’s numeracy score (wave 1) 
* `school_clus.W1`: School cluster membership (wave 1)
* `time`: Wave
* `school_clus`: School cluster membership (waves 2,3 and 4)
* `t_score`: Teacher’s  numeracy rating (waves 2,3 and 4) 
* `prev_dep`: Depressive symptoms (waves 1,2 and 3)
* `prev_sdq` : Child behaviour problems reported by SDQ (waves 1,2 and 3)

A detailed description of the variables and the data are provided in the main text.

```{r tidy=T}
#Load and look
CATS_dataL<-read.csv("CATS_dataL.csv")
head(CATS_dataL)

```

## Loading required libraries

The following packages will be required
```{r load_data,echo=TRUE, message=FALSE, warning=FALSE}
library(lme4) #for fitting lmer
library(dplyr)  ##for data wrangling
library(miceadds)  ##for FCS-3L-ml.lmer
library(mice)     ##for FCS approaches (FCS-1L and FCS-2L)
library(mitml) #for pooling and analysing the results
library(jomo)  #for JM approaches(JM-1L and JM-2L)
library(modeest)  # for finding the mode


```

### 1. Available case analysis (ACA)

An ACA  fitting a cross-classified random effects (CCREM) model can be conducted  using the lmer function in R as shown below.

```{r ACA}
lmer(t_score~ prev_dep+time+c1dage+tscore_W1+cdgender+SES+
                    (1|uniqueid)+(1|school_clus), data=CATS_dataL) 

```

### 2. First-cluster approach using single-level JM with DI for schools with repeated measures analysed in wide format (JM-1L-DI-wide_f)

The  data used in (1) above was in long format, with one row per individual per wave. For ***JM-1L-DI-wide***, we need to first reshape the data which into wide format with one row per individual. This can be done using the reshape function

```{r reshape}
  CATS_dataw <- reshape(CATS_dataL,v.names=c("school_clus","t_score","prev_sdq","prev_dep"),timevar = "time",idvar="uniqueid",direction= "wide")
```

In order to guarantee that the correct function is being used by `jomo`, make sure that continuous variables are stored as numeric vectors in the data frame and binary/categorical variables as factors. To include fully observed categorical covariates with three or more categories in the predictor matrix (X) of the imputation model, appropriate dummy variables will have to be created.

Under this approach, the incomplete variables  will need to be specified as the outcomes(Y) in the single level imputation model, while all complete variables will serve as the predictors(X). The imputation model can be specified as follows: 

```{r SLJMDIWIDE_f,results="hide"}

#Remove school cluster indicators at waves 3 and 4 (as we are only using the cluster membership at wave 2)
CATS_data_f <- select(CATS_dataw,-c(school_clus.3,school_clus.4))

##Set number of imputations, burn-ins and between-imputation iterations
M=3     #set to 3 to reduce the time of demo, 20 more appropriate
nburn=1 #set to 1 to reduce the time of demo, 1000 more appropriate
NB=1      #set to 1 to reduce the time of demo, 100 more appropriate  

##Create a dataframe with variables to be imputed
myvars <- names(CATS_data_f) %in% c("prev_dep.2","prev_dep.3","prev_dep.4","tscore_W1","t_score.2","t_score.3","t_score.4","prev_sdq.2","prev_sdq.3","prev_sdq.4")
dataimp=CATS_data_f[myvars]

##Create a dataframe with complete variables
datacomp= cbind(Intercept=rep(1,nrow(CATS_data_f)),CATS_data_f[,names(CATS_data_f)%in%c("SES","c1dage","cdgender","school_clus.2")])

##Create school cluster dummy indicators
datacomp_dummy=fastDummies::dummy_cols(datacomp, select_columns =c("school_clus.2"),remove_first_dummy = TRUE)

#Remove the orginal school cluster inidcator at wave 2 
datacomp_dummy <- datacomp_dummy %>% select(-c("school_clus.2"))


##Perform imputations without the random effects (single-level imputation)
imp1<-jomo1con(Y=dataimp, X=datacomp_dummy, nimp=M,nburn=nburn,nbetween=NB,set.seed(291002))
#output not shown  to reduce length of vignette
```

After imputation, the imputed datasets will need to be reshaped into long format and the orginal school cluster memberships (which are time-varying) will need to be reattached before fitting the substantive analysis model.

```{r analysis1}
mylist=list()

for(m in 1:M)
{
  #1.Extract the mth imputed dataset
  datw<-imp1[imp1$Imputation==m,]
  
  #2.Remove indicator variables and other unwanted variables
  datw=datw %>% select(!contains(c("school_clus.","sdq")),-Intercept,-Imputation)
  
  #3.Reattach the school variable and remove school indicator variables
  datw=cbind(datw,CATS_dataw$school_clus.2,CATS_dataw$school_clus.3,CATS_dataw$school_clus.4)
  
  #4.Remove unwanted variables
  datw=datw %>%dplyr:: rename(school_clus.2="CATS_dataw$school_clus.2",school_clus.3="CATS_dataw$school_clus.3",
                       school_clus.4="CATS_dataw$school_clus.4")
  
  #5.Reshape to long
  datL=reshape(datw,varying =list(c("prev_dep.2","prev_dep.3","prev_dep.4"),
                                  c("t_score.2","t_score.3","t_score.4"),
                                  c("school_clus.2","school_clus.3","school_clus.4")),idvar="id", 
               v.names=c("prev_dep","t_score","school_clus"), times=c(2,3,4),direction= "long")
  
  datL <- datL[order(datL$school,datL$id),]    
  
  #6.save the datasets in a list
  mylist[[m]]= datL
}


```

For fitting the analysis model on the imputed datasets and pooling the results, the `mitml’ package can be used. 

```{r fit}

#Fit the analysis of interest on the imputed datasets 
mods <- lapply(mylist,function(d) {lmer(t_score~ prev_dep+time+c1dage+tscore_W1+cdgender+SES+
                                          (1|id)+(1|school_clus), data=d) } )
testEstimates(mods,var.comp=T)

```

### 3. First-cluster approach using single-level FCS with DI for schools with repeated measures analysed in wide format (FCS-1L-DI-wide_f)

Similar to the ***JM-1L-DI-wide_f*** approach, here the dataset will need to be in wide format during the imputation process. To carry out single-level FCS, R package `mice` can be used. As all incomplete variables are continuous, the method used is `norm`. More details on methods within `mice` for different types of variables can be found [here](https://stefvanbuuren.name/fimd/). Here each of the incomplete variable will be imputed by using a series of univariate imputation models with the incomplete variable as the outcome and all the other variables as the predictors.  

```{r FCS1LDIWIDE_f}

##Set the number of imputations 
M<-3     #set to 3 to reduce the time of demo, 20 more appropriate

##Recode school cluster at wave 2 as a categorical variable
CATS_dataw$school_clus.2=as.factor(CATS_dataw$school_clus.2)

#Generate the predictor matrix
pred <- mice::make.predictorMatrix(CATS_dataw)

#In the predictor matrix, a value of
#0: Indicates that the column variable is not used as predictor for the row variable
#1: Indicates that the column variable is used as a predcitor with a fixed effect for the row variable
#2:  Indicates that the column variable is used as a predcitor with a fixed and a random effect for the row variable
#-2:  Indicates that column variable is the cluster/group variable (Only one variable is allowed)
pred[,c("school_clus.3","school_clus.4","uniqueid")] <- 0

##Perform imputations
imp2=mice(data=CATS_dataw, predictorMatrix = pred,m=M,method="norm",maxit=1)

```

Similar to ***JM-1L-DI-wide_f***, after imputation, the imputed datasets will need to be reshaped into long format before fitting the substantive analysis model. 

```{r }

mylist=list()

for(m in 1:M)
{
  #1. Extract the mth imputed dataset
  datw<-mice::complete(imp2,m)
  
  #2. Remove unwanted variables
  datw=datw %>% select(!contains("sdq"))
  
  #3. Recode school cluster at waves 3 and 4 as a categorical variable
  datw[,c("school_clus.3","school_clus.4")] <- lapply(datw[,c("school_clus.3","school_clus.4")] , factor)
  
  #4. Reshape to long
  datL=reshape(datw,varying =list(c("prev_dep.2","prev_dep.3","prev_dep.4"), c("t_score.2","t_score.3","t_score.4"),c("school_clus.2","school_clus.3","school_clus.4")),idvar="uniqueid",v.names=c("prev_dep","t_score","school_clus"), times=c(2,3,4),direction= "long")
  
  datL <- datL[order(datL$uniqueid),]    
  
  #5. Save the dataset in a list
  mylist[[m]]= datL
}

```

```{r analysis2}

#Fit the analysis of interest on the imputed datasets 
mods <- lapply(mylist,function(d) {lmer(t_score~ prev_dep+time+c1dage+tscore_W1+cdgender+SES+
                                          (1|uniqueid)+(1|school_clus), data=d) } )
testEstimates(mods,var.comp=T)

```

### 4. First-cluster approach using three-level FCS approach in Blimp  (FCS-3L-Blimp_f)

The syntax file for conducting imputations under this approach can be found [here](https://github.com/rushwije/MI3level_cross-classfied). 

**Note**: Prior to importing data to the Blimp-studio software, recode all missing values (for example as 999) and categorical variables as numerical variables. This is because special characters in the data such as NA values and string values can cause the data import to fail. The user guide, examples and the new version releases for the Blimp-studio software can be found [here](http://www.appliedmissingdata.com/multilevel-imputation.html).

The multiple imputed datasets will be saved in a single file in a stacked format.The substantive model then needs to be fitted to each imputed dataset and results pooled across the datasets. 

### 5. Common-cluster approach using single-level JM with DI for schools with repeated measures analysed in wide format (JM-1L-DI-wide_c)

The imputation procedure and the syntax under this approach is similar to ***JM-1L-DI-wide_f*** except here prior to imputations, the mode of the school cluster membership at waves 2,3 and 4 will need to be computed to be used as the school cluster membership for each individual.  

```{r SLJMDIWIDE_c,results="hide"}

CATS_datam=CATS_dataw

##Generate school cluster indicator (mode of school clus at 2,3,and 4)
School_clusters=CATS_datam %>% select(school_clus.2,school_clus.3,school_clus.4)

##Find the mode and attach it to CATS_data
CATS_datam$schoolclus_mode=apply(School_clusters,1,mfv1)

##Set number of imputations, burn-ins and the between imputation iterations
M=3     #set to 3 to reduce the time of demo, 20 more appropriate
nburn=1 #set to 1 to reduce the time of demo, 1000 more appropriate
NB=1      #set to 1 to reduce the time of demo, 100 more appropriate  

myvars <- names(CATS_datam) %in% c("prev_dep.2","prev_dep.3","prev_dep.4","tscore_W1","t_score.2","t_score.3","t_score.4","prev_sdq.2","prev_sdq.3","prev_sdq.4" ) 
dataimp=CATS_datam[myvars]

##Create a dataframe with complete variables
datacomp= cbind(Intercept=rep(1,nrow(CATS_datam)),
                CATS_datam[,names(CATS_datam)%in%c("SES","c1dage","cdgender","schoolclus_mode")])

##Create school dummy indicators
datacomp_dummy=fastDummies::dummy_cols(datacomp, select_columns =c("schoolclus_mode"),remove_first_dummy = TRUE)

#Remove the original school cluster indicators 
datacomp_dummy <- datacomp_dummy %>% select(-c("schoolclus_mode"))

##Perform imputations without the random effects (SL imputation)
imp3<-jomo1con(Y=dataimp, X=datacomp_dummy, nimp=M,nburn=nburn,nbetween=NB, set.seed(278100))
#output not shown  to reduce length of vignette

```

After imputation, the imputed datasets will need to be reshaped into long format and the orginal school cluster memberships will need to be reattached before fitting the substantive analysis model as before.

### 6. Common-cluster approach using single-level FCS with DI for schools with repeated measures analysed in wide format (FCS-1L-DI-wide_c)

The imputation procedure and the syntax under this approach is similar to ***FCS-1L-DI-wide_f*** except here the mode of the school cluster membership at waves 2,3 and 4 will be used as the school cluster membership for each individual.  

```{r SLFCSDIWIDE_C}

CATS_datamFCS <- CATS_dataw

##Generate school cluster indicator (mode of school clus at 2,3,and 4)
School_clusters=CATS_dataw %>% select(school_clus.2,school_clus.3,school_clus.4)

##Find the mode and attach it to CATS_data
CATS_datamFCS$schoolclus_mode=apply(School_clusters,1,mfv1)
CATS_datamFCS$schoolclus_mode=as.factor(CATS_datamFCS$schoolclus_mode)

##Set number of imputations and number of burn-in iterations
M<-3     #set to 3 to reduce the time of demo, 20 more appropriate

##Generate the predictor matrix
pred <- mice::make.predictorMatrix(CATS_datamFCS)

#In the predictor matrix, a value of
#0: Indicates that the column variable is not used as predictor for the row variable
#1: Indicates that the column variable is used as a predcitor with a fixed effect for the row variable
#2:  Indicates that the column variable is used as a predcitor with a fixed and a random effect for the row variable
#-2:  Indicates that column variable is the cluster/group variable (Only one variable is allowed)
pred[,c("school_clus.2","school_clus.3","school_clus.4","uniqueid")] <- 0

##Perform imputations
imp4=mice(data=CATS_datamFCS, predictorMatrix = pred,m=M,method="norm",maxit=1)

```

Post-imputation, the imputed datasets will need to be reshaped into long format and the orginal school cluster memberships will need to be reattached before fitting the substantive analysis model as before

### 7. Common-cluster approach using three-level FCS approach in Blimp  (FCS-3L-Blimp_c)

For the syntax file see [here](https://github.com/rushwije/MI3level_cross-classfied).

### 8. Single-level FCS with DI indicators for the higher-level clusters and repeated measures imputed in wide format (FCS-1L-DI-wide) 

Under this approah, each of the incomplete variables will be imputed by using a series of univariate imputation models with the incomplete variable as the outcome and all the other variables as the predictors. In each of the univariate imputation models specified for each incomplete repeated measure a set of DIs for the school cluster membership at that particular wave will be included. This is achieved using the predictor matrix as shown below. 

```{r SLFCSDIWIDE}

CATS_dataFCSv <- CATS_dataw

##Set number of imputations 
M<-3   #set to 3 to reduce the time of demo, 20 more appropriate

##Recode school cluster at wave 2 as categorical 
CATS_dataFCSv[,c("school_clus.W1","school_clus.2","school_clus.3","school_clus.4")] <- lapply(CATS_dataFCSv[,c("school_clus.W1","school_clus.2","school_clus.3","school_clus.4")] , factor)


#Generate the predictor matrix
pred <- mice::make.predictorMatrix(CATS_dataFCSv)

#In the predictor matrix, a value of
#0: Indicates that the column variable is not used as predictor for the row variable
#1: Indicates that the column variable is used as a predcitor with a fixed effect for the row variable
#2:  Indicates that the column variable is used as a predcitor with a fixed and a random effect for the row variable
#-2:  Indicates that column variable is the cluster/group variable (Only one variable is allowed)

pred[,"uniqueid"] <- 0  

#set the cluster membership for each incomplete repeated measure selectively in the predictor matrix
pred[c("prev_dep.2","t_score.2","prev_sdq.2"),c("school_clus.W1","school_clus.3","school_clus.4")]<-0  
pred[c("prev_dep.3","t_score.3","prev_sdq.3"),c("school_clus.W1","school_clus.2","school_clus.4")]<-0 
pred[c("prev_dep.4","t_score.4","prev_sdq.4"),c("school_clus.W1","school_clus.2","school_clus.3")]<-0 
pred["tscore_W1",c("school_clus.2","school_clus.3","school_clus.4")]<-0 

##Perform imputations
imp5=mice(data=CATS_dataFCSv, predictorMatrix = pred,m=M,method="norm",maxit=1)

```

Post-imputation, the imputed datasets will need to be reshaped into long format before fitting the substantive analysis model as before. 

### 9. Two-level FCS for the higher-level clusters with repeated measures imputed in wide format (FCS-2L-wide) 

Underthis approach, the data will be in wide format during the imputation stage. Two-level univariate linear imputation  models are specified for each incomplete variable. To specify the imputation models, the method `2l.pan` in the R package `mice` can be used. Here for each univariate imputation model as before an  predictor variables to be used will need to be specified. In addition, in each model, the school cluster membership will too need to be included as the cluster/group variable. This cam be done by using the integer value (-2) in the predictor matrix specification, indicating the cluster variable is to be modelled via random effects.

```{r FCS2LDIWIDE}

##set the number of imputations
M=3   #set to 3 to reduce the time of demo, 20 more appropriate


#create the predictor matrix
pred=make.predictorMatrix(CATS_dataw)

#In the predictor matrix, a value of
#0: Indicates that the column variable is not used as predictor for the row variable
#1: Indicates that the column variable is used as a predcitor with a fixed effect for the row variable
#2:  Indicates that the column variable is used as a predcitor with a fixed and a random effect for the row variable
#-2:  Indicates that column variable is the cluster/group variable (Only one variable is allowed)- needs to be coded as integer

##Recode school cluster at wave 2 back to an integer variable-as it was convereted to a factor in (3) above
CATS_dataw$school_clus.2=as.integer(CATS_dataw$school_clus.2)


#Remove variables that are not required to be imputed and other unwanted predictor variables from the predictor matrix 
pred[c("uniqueid","cdgender","c1dage","school_clus.W1","school_clus.2","school_clus.3","school_clus.4","SES"),]<-0
pred[,"uniqueid"]=0
pred[c("prev_dep.2","t_score.2","prev_sdq.2"),c("school_clus.3","school_clus.4")]<-0
pred[c("prev_dep.3","t_score.3","prev_sdq.3"),c("school_clus.2","school_clus.4")]<-0
pred[c("prev_dep.4","t_score.4","prev_sdq.4"),c("school_clus.2","school_clus.3")]<-0
pred["tscore_W1",c("school_clus.2","school_clus.3","school_clus.4")]<-0

#specify the cluster variables for each incomplete variable
pred[c("prev_dep.2","t_score.2","prev_sdq.2"),"school_clus.2"]=-2
pred[c("prev_dep.3","t_score.3","prev_sdq.3"),"school_clus.3"]=-2
pred[c("prev_dep.4","t_score.4","prev_sdq.4"),"school_clus.4"]=-2
pred["tscore_W1","school_clus.W1"]<--2

#Perform the imputations
imp6<-mice(CATS_dataw,m=M, maxit=1,predictorMatrix=pred, method="2l.pan")

```

Post-imputation, the imputed datasets will need to be reshaped into long format before fitting the substantive analysis model.

### 10. Two-level JM for repeated measures with DI for higher-level clusters (JM-2L-DI)

Under this approach, the data can be imputed in long format because the clustering between repeated measures of the same variable is now modelled using the two-level joint imputation model in `jomo`. 
The incomplete variables will be specified as outcomes(Y) in the two level imputation model, while all complete variables will serve as the predictors(X) as before. However, here we also have to define incomplete level-2 variables as well.As before to include fully observed categorical covariates with three or more categories in the predictor matrix of the imputation model, appropriate dummy variables have to be created. An additional data frame (Z) will also need to be created for the random intercepts for each individual and the cluster indicator, individual’s uniqueid, will need to be specified in the imputation model as shown below. 

```{r JM2LDI,results="hide"}
##Set number of imputations 
M=3     #set to 3 to reduce the time of demo, set to 20 
nburn=1 #set to 1 to reduce the time of demo, set to 1000 
NB=1      #set to 1 to reduce the time of demo, set to 100 

##Create a dataframe with variables to be imputed

#For level 1 variables
myvars1 <- names(CATS_dataL) %in% c("prev_dep","prev_sdq","t_score") 
dataimp1=CATS_dataL[myvars1]

#For level 2 variables
myvars2 <- names(CATS_dataL) %in% c("tscore_W1") 
dataimp2=CATS_dataL[myvars2]

##Create a dataframe with complete variables for imputing level 1 varriables
datacomp1=cbind(Intercept=rep(1,nrow(CATS_dataL)),CATS_dataL[,names(CATS_dataL)%in%c("SES","c1dage","cdgender","time","school_clus")])

#Create a dataframe with complete variables for imputing level 2 variables
datacomp2=cbind(Intercept=rep(1,nrow(CATS_dataL)),CATS_dataL[,names(CATS_dataL)%in%c("SES","c1dage","cdgender")])

##create school dummy indicators for waves, 2,3 and 4
datacomp_dummy=fastDummies::dummy_cols(datacomp1, select_columns =c("school_clus"),remove_first_dummy = TRUE)

#Remove school cluster indicator 
datacomp_dummy <- datacomp_dummy %>% select(!school_clus)

#Create a data frame with complete variables with RANDOM EFFECTS, with column of 1's for intercept
datLcompRE<-cbind(Intercept=rep(1,nrow(CATS_dataL)))

##Perform imputations with the random effects (2L imputation)
imp7<-jomo(Y=dataimp1,Y2=dataimp2,X=datacomp_dummy,X2=datacomp2,Z=datLcompRE,clus=CATS_dataL$uniqueid,nimp=M,nburn=nburn,nbetween=NB)
#output not shown  to reduce length of vignette

```

Post-imputation, no reshape is required as the data was imputed in long format. However, the original school cluster variable will be need to be reattached as it was replaced by dummy inidcators in the imputation stage. 

```{r }
mylist=list()

for(m in 1:M)
{
  #1. Extract the mth imputed variable 
   datL<-imp7[imp7$Imputation==m,]
  
  #2. Attach the original school variables
  datL=cbind(datL,CATS_dataL$school_clus)
 
  #3. Remove unwanted variables
  datL <- datL[!duplicated(as.list(datL))]
  datL=datL %>% select(!contains(c("school_clus_","prev_sdq")),-Intercept,-id)
  datL=datL %>% rename(school=`CATS_dataL$school_clus`)
  
  #4. save the dataset in a list
  mylist[[m]]= datL
}

#Fit the analysis of interest on the imputed datasets 
mods <- lapply(mylist,function(d) {lmer(t_score~ prev_dep+time+c1dage+tscore_W1+cdgender+SES+ (1|clus)+(1|school), data=d) } )

testEstimates(mods, var.comp=TRUE)

```


### 11. Two-level FCS for repeated measures with DI for higher-level clusters (FCS-2L-DI)

Similar to ***JM-2L-DI*** here the repeated measures will be imputed in long format, with the clustering between repeated measures of the same variable modelled using a univariate two-level imputation model. The imputation models can be specified using the R package `mice’. 

For each univariate imputation model, an imputation method and the predictor variables to be used will need to be specified as shown below. When specifying the predictor matrix, specify the id variable `uniquieid` as the group variable (-2) , remove school cluster membership at wave 1 from the predicor set for all variables and finally, remove the time-varying school cluster membership from the imputation model of the time-fixed baseline variable `tscore_W1`. The latter is because the imputation method used for imputing level-2 variable, `2lonly.norm`, averages all level 1 predictor to impute the level -2 variable by using a linear model. Aggregrating school cluster inidcators this way is not practical and it results in implausible imputed values. 

```{r FCS2LDI}

# Recode the cluster variable from factor to integer
CATS_dataL$uniqueid <- as.integer(CATS_dataL$uniqueid)

##set the number of imputations 
M=3 #set to 3 to reduce the time of demo, 20 more appropriate

#Recode all cluster indicator variables to factor
CATS_dataL[,c("school_clus","school_clus.W1")] <- lapply(CATS_dataL[,c("school_clus","school_clus.W1")],factor)

#Create the predictor matrix
pred=make.predictorMatrix(CATS_dataL)

#In the predictor matrix, a value of
#0: Indicates that the column variable is not used as predictor for the row variable
#1: Indicates that the column variable is used as a predcitor with a fixed effect for the row variable
#2:  Indicates that the column variable is used as a predcitor with a fixed and a random effect for the row variable
#-2:  Indicates that column variable is the cluster/group variable (Only one variable is allowed)


pred[,"uniqueid"]=-2 # ID variable to be used as the cluster variable
pred[c("uniqueid","cdgender","c1dage","time","SES","school_clus","school_clus.W1"),]<-0#Set fully observed variables to 0 as they don't need to be imputed
pred["tscore_W1","school_clus"]<-0    #Remove the school cluster variable from the imputation model for                                                #time-fixed variable (This is because aggregrating them to the                                                #individual level
pred[c("prev_dep","prev_sdq","t_score"),"school_clus.W1"] <- 0 #Remove school cluster inidcator at wave 1 from the                                                          #imputation models of time-varying variables

##Specify the imputation method
meth=make.method(CATS_dataL)
meth[c( "t_score","prev_sdq" ,"prev_dep")]="2l.pan"
#meth["tscore_W1"]="2lonly.norm"  #Here level 2 variable not imputed due to sparseness in simulated data, in practice this needs to be imputed using a method for imputing level 2 variables

#Perform the imputations
imp8<-mice(CATS_dataL,m=M, maxit=1,predictorMatrix=pred, method=meth)

```

Post-imputation, no reshape is required as the data was imputed in long format and the anlysis can be carried out on the imputed data. 

```{r }
mylist=list()

for(m in 1:M)
{
  datL<-mice::complete(imp8,m)
  
  #5. save the dataset in a list
  mylist[[m]]= datL
}

#Fit the analysis of interest on the imputed datasets 
mods <- lapply(mylist,function(d) {lmer(t_score~ prev_dep+time+c1dage+tscore_W1+cdgender+SES+
                                          (1|uniqueid)+(1|school_clus), data=d) } )
testEstimates(mods, var.comp=TRUE)

```

### 12. Three-level FCS (FCS-3L-ml.lmer)

Under this approach a univariate CCREM is specified for each incomplete variable. The imputation method `ml.lmer` from the R package `miceadds’ can be used for this As a three-level model is used, the repeated measures will be imputed in long format. As in the previous FCS approaches, a predictor matrix and an imputation method will need to be specified for each incomplete variable.   

```{r }
# create default predictor matrix and imputation methods
predMatrix <- make.predictorMatrix(data = CATS_dataL)
impMethod <- make.method(data = CATS_dataL)


# method for lower-level variables :tscore_W1 at level 2, and t_score,c_sdq and prev_sdq at level 1
impMethod["tscore_W1"] <- "ml.lmer"
impMethod["t_score"] <- "ml.lmer"
impMethod["prev_sdq"] <- "ml.lmer"
impMethod["prev_dep"] <- "ml.lmer"

# remove cluster indicator variables from set of predictors
predMatrix[, c("uniqueid", "school_clus")] <- 0

```

In addition to the predictor matrix, for variables that will be imputed using the method ml.lmer, the hierarchical structure must be set with two additional arguments (i.e. outside the predictor matrix). For all higher-level variables (i.e. those at level 2 and level 3) we need to specify the level at which the variables are measured.  For rest of the variables (i.e. for level 1 variables) the level can be left blank.

```{r }

# specify levels of higher-level variables 
level <- character(ncol(CATS_dataL))
names(level) <- colnames(CATS_dataL)
level["cdgender"] <- "uniqueid"    #for sex
level["c1dage"] <- "uniqueid"      #for age
level["tscore_w1"] <- "uniqueid"    #for teacher rating at wave 1
level["SES"] <- "uniqueid"         # for SES

```

Next for each variable being imputed, cluster variables that define the hierarchical structure in the imputation model must be specified as follows. By default, this uses a random intercept model with random effects at each of the specified levels. 

```{r }
# specify cluster indicators (as list)
cluster <- list()
cluster[["tscore_W1"]] <- c("school_clus")   #teacher rating at wave 1
cluster[["t_score"]] <- c("uniqueid","school_clus") # the outcome
cluster[["prev_sdq"]] <-  c("uniqueid","school_clus")   #the auxiliary variable
cluster[["prev_dep"]] <-  c("uniqueid","school_clus")  #the exposure

```

Then the imputation model can be specified as follows: 
```{r }
imp <- mice(CATS_dataL, method = impMethod, predictorMatrix = predMatrix, maxit = 1, 
            m = 3, levels_id = cluster, variables_levels = level,
            aggregate_automatically = FALSE)

```
Post-imputation, no reshape is required as the data was imputed in long format and the anlysis can be carried out on the imputed data. 


